An Embedding Framework for Consistent Polyhedral Surrogates

### 2019-05-23 neurips-19/poly-embed.tex
Typo fix
topk

\usetikzlibrary{calc}
\newcommand{\Comments}{1}
\newcommand{\mynote}[2]{\ifnum\Comments=1\textcolor{#1}{#2}\fi}
\newcommand{\mytodo}[2]{\ifnum\Comments=1%
  \todo[linecolor=#1!80!black,backgroundcolor=#1,bordercolor=#1!80!black]{#2}\fi}
\newcommand{\raf}[1]{\mynote{green}{[RF: #1]}}
\newcommand{\raft}[1]{\mytodo{green!20!white}{RF: #1}}
\newcommand{\jessie}[1]{\mynote{purple}{[JF: #1]}}
\newcommand{\jessiet}[1]{\mytodo{purple!20!white}{JF: #1}}
\newcommand{\bo}[1]{\mynote{blue}{[Bo: #1]}}
\newcommand{\botodo}[1]{\mytodo{blue!20!white}{[Bo: #1]}}
\newcommand{\btw}[1]{}%TURN OFF FOR NOW \mytodo{gray}{#1}}
\ifnum\Comments=1               % fix margins for todonotes
  \setlength{\marginparwidth}{1in}
\fi

@@

Surrogates for this problem commonly take reports $u\in\reals^n$, with the link $\psi(u) = \{u_{[1]},\ldots,u_{[k]}\}$, where $x_{[u]}$ is the $i^{th}$ largest component of $u$, with ties broken arbitrarily.
\citet{lapin2015top, lapin2016loss, lapin2018analysis} provide the following convex surrogate loss for this problem, which \citet{yang2018consistency} show to be inconsistent:%
\footnote{Yang and Koyejo also introduce a consistent surrogate, but it is non-convex.}
\begin{align}\label{eq:L-2-surrogate}
  L'(u)_y := \left( \frac{1}{k} \sum_{i=1}^k (u + \ones - e_y)_{[i]} - u_y \right)_+~,
\end{align}
where $e_y$ is $1$ in compnonent $y$ and 0 elsewhere.

@@

We next show that $\U = \{u \in \{0,1\}^n : \|u\|_0 \leq k\}$ is always represented among the optimizers of $L'$. \begin{lemma}\label{lem:top-k-optimal-corners}
  For all $p \in \simplex$ we have $\prop{L'}(p) \cap \U \neq \emptyset$.
\end{lemma}
\begin{proof}
  Let us first simplify $L'$:
  \[
    L'(u)_y = \left( 1 - u_y + \frac{1}{k} \sum_{i=1}^k (u - e_y)_{[i]} \right)_+~.
  \]
  Next, we observe that $L'$ is invariant in the $\ones$ direction, so without loss of generality, we can restrict our consideration to reports so that $u_{[n]} = 0$.
  To see this,
  \begin{align*}
  L'(u + \alpha \ones)_y &=\left(\frac{1}{k} \sum_{i=1}^k ((u+\alpha \ones) + \ones - e_y)_{[i]} - (u_y + \alpha) \right)_+\\
  &= \left(\frac{1}{k} \left[k\alpha + \sum_{i=1}^k (u + \ones - e_y)_{[i]}\right] - (u_y + \alpha) \right)_+\\
  &= \left(\alpha + \frac{1}{k} \left[\sum_{i=1}^k (u + \ones - e_y)_{[i]}\right] - (u_y + \alpha) \right)_+\\
  &= \left(\frac{1}{k} \left[\sum_{i=1}^k (u + \ones - e_y)_{[i]}\right] - u_y \right)_+\\
  &= L(u)_y
  \end{align*}

  We first claim that for all $y \in \Y$, the minimum of $L(u)_y$ is achieved by some $u \in [0,1]^n$.
  Observe that if $u_{[1]} > 1$, the value of the loss gets ``cropped'' and will be the same by reporting $u_{[1]} = u_{[n]} + 1 = 1$. \jessiet{Flesh this out later.}
  Since $u_{[n]} = 0 \implies u_{[1]} \in [0,1]$, we can then observe there is some $u \in \prop{L'}$ so that $u \in [0,1]^n$.

  Now, since we can restrict to the closed hypercube, we claim that we can rewrite the expected loss of $L'$ as
  \begin{align*}
  	\inprod{p}{L'(u)} &= \frac{1}{k} \left( \sum_{i=1}^k (1 - p_{[i]}) u_{[i]} + u_{[k+1]} \sum_{i=1}^k p_{[i]} \right) + 1 - \inprod{p}{u}
  \end{align*}

  To verify this, observe the first term of the summation comes from the probability that you incorrectly did not have $u_y$ in the top-$k$, and the second term comes from the $(k+1)^{th}$ term being added when $u_y$ is correctly in the top $k$ (and therefore not an averaged term, since $u_y \in [0,1]$.)
  The $1$ is pulled out from the average, and $\inprod{p}{u}$ comes from the $-u_y$ term form each $y$ by linearity of expectation.
  We don't need to concern ourselves with the max operator since $\frac{1}{k} \sum_{i=1}^k (u + \ones - e_y)_{[i]} - u_y \geq 0$ for all $y \in \Y$ on $[0,1]^n$.

>>>>

\usetikzlibrary{calc}
\newcommand{\Comments}{0}
\newcommand{\mynote}[2]{\ifnum\Comments=1\textcolor{#1}{#2}\fi}
\newcommand{\mytodo}[2]{\ifnum\Comments=1%
  \todo[linecolor=#1!80!black,backgroundcolor=#1,bordercolor=#1!80!black]{#2}\fi}
\newcommand{\raf}[1]{\mynote{green}{[RF: #1]}}
\newcommand{\raft}[1]{\mytodo{green!20!white}{RF: #1}}
\newcommand{\jessie}[1]{\mynote{purple}{[JF: #1]}}
\newcommand{\jessiet}[1]{\mytodo{purple!20!white}{JF: #1}}
\newcommand{\bo}[1]{\mynote{blue}{[Bo: #1]}}
\newcommand{\botodo}[1]{\mytodo{blue!20!white}{[Bo: #1]}}
\newcommand{\btw}[1]{}%TURN OFF FOR NOW \mytodo{gray}{#1}}
\ifnum\Comments=1               % fix margins for todonotes
  \setlength{\marginparwidth}{1in}
\fi

@@

Surrogates for this problem commonly take reports $u\in\reals^n$, with the link $\psi(u) = \{u_{[1]},\ldots,u_{[k]}\}$, where $x_{[u]}$ is the $i^{th}$ largest component of $u$, with ties broken arbitrarily.
\citet{lapin2015top, lapin2016loss, lapin2018analysis} provide the following convex surrogate loss for this problem, which \citet{yang2018consistency} show to be inconsistent:%
\footnote{Yang and Koyejo also introduce a consistent surrogate, but it is non-convex.}
\begin{align}\label{eq:L-2-surrogate}
  L'(u)_y := \left( \frac{1}{k} \sum_{i=1}^k (u + \ones - e_y)_{[i]} - u_y \right)_+~,
\end{align}
where $e_y$ is $1$ in component $y$ and 0 elsewhere.

@@

We next show that $\U = \{u \in \{0,1\}^n : \|u\|_0 \leq k\}$ is always represented among the optimizers of $L'$. \begin{lemma}\label{lem:top-k-optimal-corners}
  For all $p \in \simplex$ we have $\prop{L'}(p) \cap \U \neq \emptyset$.
\end{lemma}
\begin{proof}
  Let us first simplify $L'$:
  \[
    L'(u)_y = \left( 1 - u_y + \frac{1}{k} \sum_{i=1}^k (u - e_y)_{[i]} \right)_+~.
  \]
  Next, we observe that $L'$ is invariant in the $\ones$ direction, so without loss of generality, we can restrict our consideration to reports so that $u_{[n]} = 0$.
  To see this,
  \begin{align*}
    L'(u + \alpha \ones)_y &= \left( 1 - (u_y + \alpha) + \frac{1}{k} \sum_{i=1}^k ((u + \alpha) - e_y)_{[i]} \right)_+\\
    &= \left( 1 - (u_y + \alpha) + \frac{1}{k}\left[ \alpha k + \sum_{i=1}^k (u - e_y)_{[i]} \right] \right)_+\\
    &= \left( 1 -u_y + \frac{1}{k}\left[ \sum_{i=1}^k (u - e_y)_{[i]} \right] \right)_+\\
  &= L'(u)_y
  \end{align*}

  We now claim that for all $y \in \Y$, the minimum of $L(u)_y$ is achieved by some $u \in [0,1]^n$.
  This follows from the observation that if $u_{[1]} > 1$, the the loss only decreases by reporting $u_{[1]} = u_{[n]} + 1 = 1$.
  For $u\in[0,1]^n$, the argument to $(\cdot)_+$ is nonnegative, so we can rewrite the expected loss of $L'$ as
  \begin{align*}
  	\inprod{p}{L'(u)} &= \frac{1}{k} \left( \sum_{i=1}^k (1 - p_{[i]}) u_{[i]} + u_{[k+1]} \sum_{i=1}^k p_{[i]} \right) + 1 - \inprod{p}{u}
  \end{align*}
  To verify this, observe the first term of the summation comes from the probability that you incorrectly did not have $u_y$ in the top-$k$, and the second term comes from the $(k+1)^{th}$ term being added when $u_y$ is correctly in the top $k$ (and therefore not an averaged term, since $u_y \in [0,1]$.)
  The $1$ is pulled out from the average, and $\inprod{p}{u}$ comes from the $-u_y$ term form each $y$ by linearity of expectation.
